---
title: "bka_analysis"
output: html_document
---

``` {r,include = FALSE}

source('/Users/dancrowley/BKA.PREEMPT/R/soft.max.clean.R')
file_path = '~/Box Sync/Research NSF Bat1Health Collaboration/bat_immunoassays/'
file.list <- list.files(file_path, pattern = '.xlsx')
file.list


for (i in 1:length(file.list))
{
  file <- paste(file_path, file.list[i], sep = "")
  assign(file.list[i], soft.max.clean(file_path = file, num_of_time_points = 7))
}

dfs <- Filter(function(x) is(x, "data.frame"), mget(ls()))
bka.merge <- do.call(rbind, dfs)
rm(list=(ls()[ls()!="bka.merge"]))

bka.merge$sample <- tolower(bka.merge$sample)
bka.merge$sample <- gsub(" ", "", bka.merge$sample)

bka.merge$measure <- as.numeric(bka.merge$measure)

```

#BKA Results

##Bacteria Only
###Data Untransformed
``` {r,echo = FALSE}
bka.merge %>%
  filter(sample == 'bacteriaonly') %>%
  unite(well, c(which_row, column)) %>%
  ggplot(aes(x=time, y= (as.numeric(measure)))) +
  #ggplot(aes(x=time, y= log(as.numeric(measure)))) +
  #ggplot(aes(x=time, y= boot::inv.logit(as.numeric(measure)))) +
  geom_point(aes(group = well, col = bacteria)) +
  geom_line(aes(group = well, col = bacteria)) +
  #ylim(.9,2)+
  #xlim(0,.6)+
  facet_wrap(~experiment_id)

```

### Log Transformed
``` {r,echo = FALSE}
bka.merge %>%
  filter(sample == 'bacteriaonly') %>%
  unite(well, c(which_row, column)) %>%
  #ggplot(aes(x=time, y= (as.numeric(measure)))) +
  ggplot(aes(x=time, y= log(as.numeric(measure)))) +
  #ggplot(aes(x=time, y= boot::inv.logit(as.numeric(measure)))) +
  geom_point(aes(group = well, col = bacteria)) +
  geom_line(aes(group = well, col = bacteria)) +
  #ylim(.9,2)+
  #xlim(0,.6)+
  facet_wrap(~experiment_id)

```

### inverse Logit Transformed
``` {r,echo = FALSE}
bka.merge %>%
  filter(sample == 'bacteriaonly') %>%
  unite(well, c(which_row, column)) %>%
  #ggplot(aes(x=time, y= (as.numeric(measure)))) +
  #ggplot(aes(x=time, y= log(as.numeric(measure)))) +
  ggplot(aes(x=time, y= boot::inv.logit(as.numeric(measure)))) +
  geom_point(aes(group = well, col = bacteria)) +
  geom_line(aes(group = well, col = bacteria)) +
  #ylim(.9,2)+
  #xlim(0,.6)+
  facet_wrap(~experiment_id)

```

##Repliace Variance As A Function of Time

###Bacteria Only  (includes staph and e.coli)
``` {r,echo = FALSE}

bka.merge %>%
  mutate(sample = ifelse(grepl('bat',sample),'bat serum', sample)) %>%
  filter(sample == 'bacteriaonly') %>%
  #filter(bacteria == 'E.coli') %>%
  unite(well, c(which_row, column)) %>%
  unite(experiment_id, c(sample, experiment_id)) %>%
  group_by(time, experiment_id, bacteria) %>%
  summarise(variance = var(measure)) %>%
  ggplot() +
  geom_line(aes(x = time, y = variance, group = experiment_id, col= experiment_id)) #+
  #theme(legend.position="none") #+
  #facet_wrap(~sample, scales = 'free', ncol= 2)
```

##All Other Samples

Each color is a unique sample type and experiment
The Y axis is the variance for the replicates. 
My conclusion: highest variance at the first reading and the final readings. 
We see this in the time series data in the bacteria only samples, however this suggests the same pattern exists for all sample types. 

###E.Coli First

``` {r,echo = FALSE}

bka.merge %>%
  mutate(sample = ifelse(grepl('bat',sample),'bat serum', sample)) %>%
  filter(sample != 'blank') %>%
  filter(bacteria == 'E.coli') %>%
  filter(sample != 'bacteriaonly') %>%
  unite(well, c(which_row, column)) %>%
  unite(experiment_id, c(sample, experiment_id)) %>%
  group_by(time, experiment_id, bacteria) %>%
  summarise(variance = var(measure)) %>%
  ggplot() +
  geom_line(aes(x = time, y = variance, group = experiment_id, col= experiment_id)) +
  theme(legend.position="none") +
  geom_smooth(aes(x = time, y = variance), method = 'loess') #+

  #facet_wrap(~sample, scales = 'free', ncol= 2)
```

##Staph Next
``` {r,echo = FALSE}

bka.merge %>%
  mutate(sample = ifelse(grepl('bat',sample),'bat serum', sample)) %>%
  filter(sample != 'blank') %>%
  filter(bacteria == 'S.aureus') %>%
  filter(sample != 'bacteriaonly') %>%
  unite(well, c(which_row, column)) %>%
  unite(experiment_id, c(sample, experiment_id)) %>%
  group_by(time, experiment_id, bacteria) %>%
  summarise(variance = var(measure)) %>%
  ggplot() +
  geom_line(aes(x = time, y = variance, group = experiment_id, col= experiment_id)) +
  theme(legend.position="none") +
  geom_smooth(aes(x = time, y = variance), method = 'loess') #+

  #facet_wrap(~sample, scales = 'free', ncol= 2)
```




##Box Plot of Samples
### E. Coli Samples Only

``` {r,echo = FALSE}

tp12 <- bka.merge[bka.merge$time == 0.5, ]
tp12 <- tp12[tp12$bacteria == 'E.coli', ]
tp12 <- tp12[tp12$sample %in% c('bacteriaonly','batserum', 'mouseserum'), ]

tp12 %>% 
  ggplot(aes( experiment_id, as.numeric(measure))) +
  geom_hline(yintercept = 0 , color = 'red', size = 1, alpha = .5) +
  geom_boxplot() +
  geom_jitter(alpha =.5) +
  facet_wrap(~sample)  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


##Box Plot of Samples
###Staph. Aureus Samples Only

``` {r,echo = FALSE}

tp12 <- bka.merge[bka.merge$time == 0.5, ]
tp12 <- tp12[tp12$bacteria == 'S.aureus', ]
tp12 <- tp12[tp12$sample %in% c('bacteriaonly','batserum', 'mouseserum'), ]

tp12 %>% 
  ggplot(aes( experiment_id, as.numeric(measure))) +
  geom_hline(yintercept = 0 , color = 'red', size = 1, alpha = .5) +
  geom_boxplot() +
  geom_jitter(alpha =.5) +
  facet_wrap(~sample)  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


#What is an outlier? 
## Look at one sampling point only, lets choose 6 hours 
``` {r,echo = FALSE}

t.6 <- bka.merge %>%
  filter(time == .25) 

t.6 <- t.6 %>%
  unite(well, c(which_row, column)) %>%
  group_by(sample, bacteria, experiment_id) %>%
  summarise(mean = mean(measure), var = var(measure))
  

hist(t.6$var, breaks = 50)
mean(t.6$var)
var(t.6$var)

x <- t.6 %>%
  filter( var  >.1)

bka.merge %>%
  filter(sample == x$sample) %>%
  filter(bacteria == x$bacteria) %>%
  filter(experiment_id == x$experiment_id) %>%
  unite(well, c(which_row, column)) %>%
  ggplot(aes(x=time, y= (as.numeric(measure)))) +
  #ggplot(aes(x=time, y= log(as.numeric(measure)))) +
  #ggplot(aes(x=time, y= boot::inv.logit(as.numeric(measure)))) +
  geom_point(aes(group = well, col = bacteria)) +
  geom_line(aes(group = well, col = bacteria)) +
  #ylim(.9,2)+
  #xlim(0,.6)+
  facet_wrap(~experiment_id)

```


#Correlation of Time Point Delta Ratios
```{r, echo = FALSE}
source('/Users/dancrowley/BKA.PREEMPT/R/control.ratio.data.R')
ratio <- control.ratio.data(bka.merge, positive.control = 'bacteriaonly')

#remove the samples in X, we know those contain samples with a huge amount of variance

ratio <- ratio %>%
  filter(sample != x$sample) %>%
  filter(bacteria != x$bacteria) %>%
  filter(experiment_id != x$experiment_id) 

df <- ratio %>% 
  unite(well, c(which_row, column)) %>%
  filter(sample != 'bacteriaonly') %>%
  ungroup() %>%
  select(sample, well, bacteria, time.t1_12, deltaratio, experiment_id) %>%
  filter(!is.na(deltaratio)) %>%
  group_by(sample, well, bacteria, experiment_id, time.t1_12) %>%
  summarise(deltaratio= mean(deltaratio)) %>%
  tidyr::spread(time.t1_12, deltaratio)

df1 <- df[,c(5:10)]
colnames(df1) <- seq(1,6,by =1)
cor <- cor(df1, use="complete.obs")
lattice::levelplot((cor))

```

#Covariance of Time Point Delta Ratios
```{r, echo = FALSE}

cov <- cov(df1, use="complete.obs")
lattice::levelplot((cov))
```



